Improving the reliability of cognitive task measures: \break A narrative review

### 2023-01-13 manuscript/main.tex
fix typos; update to 'in press' status

% set up the author block
\setlength{\affilsep}{2em}   % set the space between author and affiliation
\author[1,*]{Samuel Zorowitz}
\author[1,2,*]{Yael Niv}
\affil[1]{Princeton Neuroscience Institute, Princeton University, USA}
\affil[2]{Department of Psychology, Princeton University, USA}
\affil[*]{Corresponding authors (zorowitz@princeton.edu, yael@princeton.edu)}

@@

As a final point, traditional sum or mean score estimates of performance (e.g., proportion correct responses, mean response time) may substantially underestimate task reliability \cite{Rouder2019-am, haines2020learning, chen2021trial}. \hl{This is because such summary scores are contaminated by trial-level noise that, in the absence of a sufficiently (possibly prohibitively) large number of trials, increases measurement error (and thus diminishes reliability). Instead, it may be preferable to use trial-level hierarchical models in which observations are organized hierarchically (e.g., individuals within a group, trials within an individual) with variability modeled at both levels. Hierarchical models exert a pooling or regularization effect on person-level variables, in effect correcting for measurement error and improving estimates of reliability} \cite{Rouder2019-am, haines2020learning, chen2021trial}. The benefits of hierarchical models for estimating reliability has been multiply demonstrated \cite{snijder2022psychometric, sullivan2022enhancing, brown2020improving, waltmann2022sufficient}, \hl{though see} \cite{rouder2019most} \hl{for discussion of when these benefits may be limited. Using statistical models that more accurately characterize the latent data-generating process (e.g., using the shifted log-normal distribution to model response times) may also improve reliability estimates} \cite{haines2020learning, price2019computational}. For a detailed discussion of hierarchical and generative models in the context of task reliability, see Haines and colleagues (this issue; \cite{haines2020learning}).

>>>>

% set up the author block
\setlength{\affilsep}{2em}   % set the space between author and affiliation
\author[1,*]{Samuel Zorowitz}
\author[1,2]{Yael Niv}
\affil[1]{Princeton Neuroscience Institute, Princeton University, USA}
\affil[2]{Department of Psychology, Princeton University, USA}
\affil[*]{Corresponding author (zorowitz@princeton.edu)}

@@

As a final point, traditional sum or mean score estimates of performance (e.g., proportion correct responses, mean response time) may substantially underestimate task reliability \cite{Rouder2019-am, haines2020learning, chen2021trial}. \hl{This is because such summary scores are contaminated by trial-level noise that, in the absence of a sufficiently (possibly prohibitively) large number of trials, increases measurement error (and thus diminishes reliability). Instead, it may be preferable to use trial-level hierarchical models in which observations are organized hierarchically (e.g., individuals within a group, trials within an individual) with variability modeled at both levels. Hierarchical models exert a pooling or regularization effect on person-level variables, in effect correcting for measurement error and improving estimates of reliability} \cite{Rouder2019-am, haines2020learning, chen2021trial}. The benefits of hierarchical models for estimating reliability have been multiply demonstrated \cite{snijder2022psychometric, sullivan2022enhancing, brown2020improving, waltmann2022sufficient}, \hl{though see} \cite{rouder2019most} \hl{for discussion of when these benefits may be limited. Using statistical models that more accurately characterize the latent data-generating process (e.g., using the shifted log-normal distribution to model response times) may also improve reliability estimates} \cite{haines2020learning, price2019computational}. For a detailed discussion of hierarchical and generative models in the context of task reliability, see Haines and colleagues (this issue; \cite{haines2020learning}).
